Software Citation Principles

### 2016-03-03 sc-principles/software-citation-principles.tex
adding new refs for motivation, shortening old motivation text and adding next text for these refs, fixing affils in appendix

\katznote{to add cite of Matthew Mayernik's similar work} 
Software and other digital resources currently appear in publications in very inconsistent ways.
For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}. Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered. 
%Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format. 
Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses. 
In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero).
% In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is machine actionable, globally unique, interoperable, and widely used by a community.
\item \textbf{Persistence}:
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}:
Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available. Should we say anything about environments?}
\item \textbf{Specificity}:
Software citations should facilitate identification of and access to the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}
\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}

@@

\katznote{should make affiliation styles more consistent}

\katznote{should check to make sure these are all still right, such as for me and Bill Mills}

@@

Carl Boettiger, UC Berkeley

@@

Melissa Haendel, OHSU

Stephanie Hagstrom, Force11

Robert Hanisch, NIST/ODI

@@

Ivan Herman, W3C

@@

Matthew B. Jones, NCEAS, UC Santa Barbara

@@

Daniel S. Katz, University of Chicago \& Argonne National Laboratory (co-chair)

@@

Ashley Sands, UCLA Information Studies

>>>>

Software and other digital resources currently appear in publications in very inconsistent ways.
For example, a random sample of 90 articles in the biology literature found seven different ways that software was mentioned, including simple names in the full-text, URLs in footnote, and different kinds of mentions in references lists: project names or websites, user manuals, publications that describe or introduce the software~\cite{howison2015jasist}. Table \ref{tab:mentions} shows examples of these varied forms of software mentions and the frequency with which they were encountered. 
%Even in the top journals (top 10 by impact factor), only 38\% of mentions included an entry of some sort in the reference list and fully 36\% were informal, following no recognizable style or format. 
Many of these kinds of mentions fail to perform the functions needed of citations, and their very diversity and frequent informality undermines the integration of software work into bibliometrics and other analyses. 
%In some ways the situation is not too surprising: practices are diverse and informal because authors receive little guidance from editors, style guides, and their scientific communities, as well as weak support from citation software (such as Endnote and Zotero). 
Studies on data and facility citation have shown similar results~\cite{parsons_duerr_minster, mayernik_poster}.
% In Howison and Bullard's study only 24\% of journals provided any guidance and software is treated inconsistently by style guides, if it appears at all.

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is machine actionable, globally unique, interoperable, and widely used by a community.
\item \textbf{Persistence}:
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}:
Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available. Should we say anything about environments?}
\item \textbf{Specificity}:
Software citations should facilitate identification of and access to the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}
\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}

@@

\katznote{should check to make sure these are all still right}

@@

Carl Boettiger, University of California,  Berkeley

@@

Melissa Haendel, Oregon Health and Science University

Stephanie Hagstrom, FORCE11

Robert Hanisch, National Institute of Standards and Technology, One Degree Imager

@@

Ivan Herman, World Wide Web Consortium (W3C)

@@

Matthew B. Jones, NCEAS, University of California,  Santa Barbara

@@

Daniel S. Katz, University of Illinois (co-chair)

@@

Ashley Sands, University of California, Los Angeles
